---
title: "Project1"
author: "Ari Boyarsky, Isaac Gritz, Clare Lohrmann, Sameer Rau, Abdul Sheikhnureldin"
date: "October 3, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Setup:

```{r, echo=TRUE}
#clean workspace
rm(list = ls())

#require igraph
library(igraph)

#read edges csv for nodeid 0
graph.edges <- read.csv(file = 'facebook/0.edges', sep=" ", header=FALSE, col.names=c("source", "target"))

#get edges
graph.e <- graph.edges - min(graph.edges) + 1
#get nodes
graph.n <- data.frame(id=seq(max(graph.e) - min(graph.e) + 1))

graph.n$group <- 1

#import into igraph
g = graph.data.frame(graph.e, directed=FALSE, vertices=graph.n)

#simplify code
g2 = simplify(g)

#plot
plot(g2)
```

# Question 3:

**Explore with some functions from ppt**

```{r, echo=TRUE}

#get verticies
V(g2)

#get edges
edges(g2)

#find vertex attributes, did not load attributes so we should not expect much here
vertex_attr(g2)

#check if the graph is simple (does not have loops or multiple edges between verticies)
is.simple(g2)


degree(g2, "198")

degree(g2)
astrocollab <- upgrade_graph(g)
plot(astrocollab)
```

# Question 4 

**Explore 10 functions**

The fucntion `eccentricity()` calulcates the shortest path distance from the farthest other node in the graph, for each node. 
```{r, echo=TRUE}

#find eccentricity of verticies, we are using head here so the markdown file does not get too long
head(eccentricity(g2))
```
The function `dfs()`, preforms a depth-first search from the root node, in this case 1. This is an algorithm to traverse the graph. `bfs()` is another function that employs an algorthim, breadth-first search, this algorithm attempts to go to each edge.
```{r, echo=TRUE}
#dfs with node 1 as root
dfs(g2, 1)

#bfs with node 1 as root
bfs(g2, 1)
```

`count_triangles()` attempts to count how many triangles a node is a part of. A triangle, is a part of a graph that has a cycle with 3 nodes and 3 edges.
```{r, echo=TRUE}

#count triangles a vertex is a part of
count_triangles(g2)
```

`is.directed()` checks to see if a graph is directed. `is.weighted()` checks to see if a graph has weights.
`is.connected()` checks to see if a graph is connected (i.e. has a path between every pair of vertices).

```{r, echo=TRUE}
#is the graph directed
is.directed(g2)

#is the graph weighted
is.weighted(g2)

#is the graph connected
is.connected(g2)
```

`head()` retrieves the first few results of a fucntion. `closeness()` computes the closeness cnetrality, how many steps to get to every other node from another node. 
```{r, echo=TRUE}
#get closeness centrality of vertices, we are using head here so the markdown file does not get too long
head(closeness(g2))
```
`count_multiple()` counts the multiple edges a node has, edges that have the same head and tail vertices.

```{r, echo=TRUE}
#find loop edges, we are using head here so the markdown file does not get too long
head(count_multiple(g2))
```

`components()` finds connected components of a graph.
```{r, echo=TRUE} 
#fine components of graph
components(g2)
```
# Question 5

```{r, echo=TRUE}
#determine central person
ac1 = alpha_centrality(g)
tail(sort(ac1),5)

#longest path, distances creates the distance table, which.max picks out highest non-inf value
which.max(distances(g2))


#largest clique, we are using head here so the markdown file does not get too long
head(largest.cliques(g))
# Hence, the largest click is 15

#ego, we are using head here so the markdown file does not get too long
head(ego(g2, gorder(g2), V(g2)))


# power centrality, we are using head here so the markdown file does not get too long
head(power_centrality(g, nodes = V(g)))
```


